---
title: "R_data handling"
author: "Lilian"
date: "7/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading Tabular Data

-. tabular data: read.table, read.csv  
-. lines of text file: readLines
-. R code files: source, dget
-. workspaces: load

### read.table
-. file = name of file  
-. header = if the first line contains the variable names/labels  
-. sep = columns separated by commas or tabs etc  
-. colClasses = numerical, character, etc  
-. comment.char = default is #, ignore anything to the right  
-. skip = skipping headers or non-data region  
-. stringsAsFactors = default is true, set to false if you dont mean to read this in as a factor  
-. for larger datasets, helps to specify as many as these as possible to reduce memory usage    
-. memory calculation = 8 bits per byte of numerical data, 2^20 bytes per MB, allow 2x  


# Textual data formats can be useful for metadata
-. dput writes R code to reconstruct an object
-. dget to read it into R
-. dump can be used on multiple objects --> source


# Subsetting Lists
-. [ returns the same object class  
-. [[ extracts elements of a list or df, but it wil not come back in the same format  
-. $ select based on name, does partially mapping    


```{r subset}
x <- c("a", "b", "c", "c", "d", "a")
x[1]
x[x > "a"] ##want everything greater than a
u <- x > "a" ##T/F if a value is greater than a
u
x[u]
```

### subsetting lists
in this example, x is a list with 2 elements: foo and bar
```{r subset list}
x <- list(foo = 1:4, bar = 0.6)
x[1]
```

single bracket returns a list, but if using a double bracket that returns a sequence (same as if i used x$foo)
```{r subset list double}
x[[1]]
```

use the [[ operator for computed indices, since dollar can only be used for the literal name. dollar.name and dollar.foo will fail here
```{r subset computed}
name <- "foo"
x[[name]]
```

to extract multiple elements of a list:
```{r subset multiple}
x[c(1,2)]
```

for nested elements:
```{r nested}
x <- list(a = list(10,12,14), b= c(3.14, 2.81))
x[[c(1,3)]]
```
this extracts 14: 3rd element of 1st object


### Subsetting Matrices
Subsetting a single column or row will give you a vector, NOT a matrix

### Removing NA values
Create a logical vector "bad" that contains NA, then remove it
```{r NA}
x <- c(1, 2, NA, 4, NA, 5)
bad <- is.na(x)
bad
x[!bad]
```

Use complete.cases in instances where you are looking for completeness in multiple objects
```{r complete}
x <- c(1,2,NA,4,NA,5)
y <- c("a","b",NA,"d",NA,"f")
good <- complete.cases(x,y)
good
x[good]
y[good]
```


# Vectorized operations
### >= greater than or equal to
### == is equal to
### %*% for true matrix multiplication


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
