---
title: "R Week 1 Quiz"
author: "Lilian"
date: "7/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solutions to problem set using hw1_data 

First, I used RStudio to import the .csv data (I had to upload this from my local directory to kgen using scp). This required installation of the readr package, followed by read_csv()

```{r load}
library(readr)
hw1_data <- read.csv("/gfs/work/llam/R-lessons/datasciencecoursera/hw1_data.csv")
```

11) In the dataset provided for this Quiz, what are the column names of the dataset?

```{r 11}
colnames(hw1_data)
```


12) Extract the first 2 rows of the data frame and print them to the console.
```{r 12}
hw1_data[1:2,]
```


13) How many observations (i.e. rows) are in this data frame?
```{r 13}
nrow(hw1_data)
```


14) Extract the last 2 rows of the data frame and print them to the console. What does the output look like?
```{r 14}
hw1_data[152:153, ]
tail(hw1_data, n=2)
```


15) What is the value of Ozone in the 47th row?
```{r 15}
hw1_data$Ozone[47]
```


16) How many missing values are in the Ozone column of this data frame?
```{r 16}
is.na(hw1_data$Ozone)
```
then sum these to get the count of NA = TRUE
```{r 16b}
sum(is.na(hw1_data$Ozone))
```


17) What is the mean of the Ozone column in this dataset? Exclude missing values (coded as NA) from this calculation.

We need to remove NA by na.rm = TRUE
```{r 17}
mean(hw1_data$Ozone, na.rm = TRUE)
```


18) Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90. What is the mean of Solar.R in this subset?
```{r 18}
mean(hw1_data[which(hw1_data$Ozone > 31 & hw1_data$Temp > 90),]$Solar.R)
```


19) What is the mean of "Temp" when "Month" is equal to 6?
```{r 19}
mean(hw1_data[which(hw1_data$Month == 6),]$Temp)
```


20) What was the maximum ozone value in the month of May (i.e. Month is equal to 5)?
```{r 20}
max(hw1_data[which(hw1_data$Month == 5),]$Ozone, na.rm = TRUE)
```




